---
title: "Social Networks and System Recomendations"
subtitle: "Project 1: "
author: "Bartosz Rożek"
date: last-modified
date-format: "DD-MM-YYYY"
editor: visual
title-block-banner: "#e8e6ef"
page-layout: article
css: Styles/styles.css
format: 
  html:
    toc: true
    smooth-scroll: true
    embed-resources: true
    fontcolor: black
    toc-location: right
    toc-title: Summary
    toc-depth: 3
    number-sections: true
    number-depth: 4
    code-fold: false
    code-summary: "Code"
tbl-cap-location: bottom
---

# Introduction

All solution are based on creating adjacency matrix using R base package options and creating the graphs with **igraph** package.

```{r message=F}
library(igraph)
```

# Exercise 1

Draw square lattice.

```{r}

lattice_graph <- function(nrow = 4, ncol = 5, quiet = F) {
    n_elements <- nrow * ncol
    
    am <- matrix(data = 0,
                 nrow = n_elements,
                 ncol = n_elements)
    
    for (i in 1:(n_elements)) {
        if (i %% ncol != 0) am[i,i + 1] <- 1
        if (i + ncol <= n_elements) am[i,i + ncol] <- 1
    }
    
    graph <- graph_from_adjacency_matrix(am, mode = "undirected")
    V(graph)$label <- ""
    
    if (!quiet) {
        graph |> plot()
      }
    return(graph)
}

lg <- lattice_graph(5, 8)
```

# Exercise 2

Draw chain. This is the same function as in the previous exercise, but one of the arguments is equal to one.

```{r}

chain <- lattice_graph(1, 11)
```

# Exercise 3

Draw complete graph.

```{r}

complete_graph <- function(n = 5, quiet = F) {
    am <- matrix(data = 1,
                 nrow = n,
                 ncol = n) - diag(n)
    graph <- graph_from_adjacency_matrix(am, mode = "undirected")
    V(graph)$label <- ""
    
    if (!quiet) {
        graph |> plot()
    }
    return(graph)
}

cg <- complete_graph(9)
```

# Exercise 4

Draw weighted graph.

```{r}

random_weighted_graph <- function(n = 5, quiet = F) {
    am <- matrix(data = runif(n ** 2, 0, 5),
                 nrow = n,
                 ncol = n)
    diag(am) <- 0
    
    graph <- graph_from_adjacency_matrix(am, weighted = TRUE, mode = "undirected")
    E(graph)$width <- E(graph)$weight + min(E(graph)$weight)
    V(graph)$label <- ""
    
    if (!quiet) {
        graph |> plot()
    }
    return(graph)
}

rwg <- random_weighted_graph(7)
```

Solution based on @ddiez answer for [stackoverflow question](https://stackoverflow.com/questions/21300821/igraph-creating-a-weighted-adjacency-matrix).

# Exercise 5

Draw directed graph.

```{r}

random_directed_graph <- function(n = 5, quiet = F) {
    am <- matrix(data = runif(n ** 2, 0, 1),
                 nrow = n,
                 ncol = n)
    diag(am) <- 0
    am[am > 0.7] <- 1
    am[am != 1] <- 0 
    graph <- graph_from_adjacency_matrix(am, mode = "directed")
    V(graph)$label <- ""
    
    if (!quiet) {
        graph |> plot()
    }
    return(graph)
}

rdg <- random_directed_graph(6)
```
