---
title: "Social Networks and System Recomendations"
subtitle: "Project 2: "
author: "Bartosz Rożek"
date: last-modified
date-format: "DD-MM-YYYY"
editor: visual
title-block-banner: "#e8e6ef"
page-layout: article
css: ../Styles/styles.css
format: 
  html:
    toc: true
    smooth-scroll: true
    embed-resources: true
    fontcolor: black
    toc-location: right
    toc-title: Summary
    toc-depth: 3
    number-sections: true
    number-depth: 4
    code-fold: false
    code-summary: "Code"
tbl-cap-location: bottom
---

# Introduction

All solution are based on creating adjacency matrix using R base package options and creating the graphs with **igraph** package.

```{r message=F}
library(igraph)
library(echarts4r)
```

# Data 1 - Game of thrones

Place vertices on a circle, in the order of their vertex ids.


```{r}


data <- read.csv("asoiaf-book1-edges.csv")
g <- graph_from_data_frame(data, directed = F, vertices = NULL)
nodes_names <- table(data$Source)
V(g)$label <-
    ifelse(names(V(g)) %in% names(nodes_names)[nodes_names > 20], names(V(g)), "")
V(g)$label <- ""
deg <- degree(g, mode = "all")
lay <- layout_in_circle(g)
V(g)$label.cex = 1
nodes_names[nodes_names > 10] |> sort(decreasing = T) |> print()
hist(nodes_names) |> plot()
plot(g, vertex.size = deg / 10, layout = lay)


```

```{r}
nodes_names <- table(data$Source)
edges <- data
nodes <- data.frame(
    name = names(nodes_names),
    value = as.numeric(nodes_names),
    size = as.numeric(nodes_names),
    stringsAsFactors = FALSE
)

g <- e_charts() |> 
    e_graph_gl() |> 
    e_graph_nodes(nodes = nodes, names = name, size = size, value = value) |> 
    e_graph_edges(edges = data, source = Source, target = Target) |> 
    e_modularity() |> 
    e_tooltip()

g


```

A circle graph makes it hard to seen, but there can be observed some vertices that "connects" groups together. 

# Data 2 - Media data

A force-directed layout algorithm, that scales relatively well to large graphs.


```{r}
nodes <- read.csv("Data files/Dataset1-Media-Example-NODES.csv")
edges <- read.csv("Data files/Dataset1-Media-Example-EDGES.csv")

g <- graph_from_data_frame(d=edges, vertices=nodes, directed = T) 
l <- layout_with_graphopt(g)
deg <- degree(g)
E(g)$width <- (E(g)$weight + min(E(g)$weight))/10
V(g)$label.cex = 0.7
plot(g, layout=l, vertex.size = deg, vertex.label=V(g)$media,
     edge.arrow.size=.4, edge.curved=0.5)

```
```{r}
edges <- merge.data.frame(edges, nodes[, c("id", "media")], by.x = "from", by.y = "id")
colnames(edges)[length(edges)] <- "from_media"
edges <- merge.data.frame(edges, nodes[, c("id", "media")], by.x = "to", by.y = "id")
colnames(edges)[length(edges)] <- "to_media"
nodes$audience.size <- nodes$audience.size
g <- e_charts() |> 
    e_graph() |> 
    e_graph_nodes(nodes = nodes, names = media, size = audience.size, 
                  value = audience.size, category = media.type) |> 
    e_graph_edges(edges = edges, source = from_media, target = to_media) |> 
    e_tooltip()

g

```

Because of the weights we can distinguish groups of news providers that often refers to each other.

# Data 3 - Airlines data

Place vertices on the plane using the force-directed layout algorithm by Fruchterman and Reingold.

```{r}
nodes <- read.csv("Data files/Dataset3-Airlines-NODES.csv")
edges <- read.csv("Data files/Dataset3-Airlines-EDGES.csv")
edges <-
    merge.data.frame(edges, nodes[, c("ID", "Code")], by.x = "Source", by.y = "ID")
colnames(edges)[length(edges)] <- "Source_media"
edges <-
    merge.data.frame(edges, nodes[, c("ID", "Code")], by.x = "Target", by.y = "ID")
colnames(edges)[length(edges)] <- "Target_media"

g <- graph_from_data_frame(d = edges,
                           vertices = nodes,
                           directed = F)
l <- layout_with_fr(g)
nodes_names <- table(edges$Target_media)
V(g)$label <- ""
deg <- degree(g)
E(g)$width <- edges$Freq / 200
V(g)$label.cex = 0.1
nodes_names[nodes_names > 10] |> sort(decreasing = T) |> print()
plot(
    g,
    layout = l,
    vertex.size = deg / 10,
    vertex.label = V(g)$media,
    edge.arrow.size = .4,
    edge.curved = 0.5
)

```
```{r}

g <- e_charts() |> 
    e_graph() |> 
    e_graph_nodes(nodes = nodes, names = Code, size = Visits, 
                  value = Visits) |> 
    e_graph_edges(edges = edges, source = Source_media, target = Target_media) |> 
    e_modularity() |> 
    e_tooltip()

g

```

It is not easy to interpret this graph because of the large number of vertices, but few things may be observed:
- there are some airports that are connecting a great amount of smaller airports
- there are airports that have only one connection to the "knot" vertex


# Data 4 - Bio diseasome

DrL is a force-directed graph layout toolbox focused on real-world large-scale graphs, developed by Shawn Martin and colleagues at Sandia National Laboratories.

```{r}
nodes <- read.csv("Data files/bio-diseasome.txt", sep = " ")
g <- graph_from_data_frame(d = nodes,
                           directed = F)
l <- layout_with_drl(g)
deg <- degree(g)
V(g)$label <- ""

plot(
    g,
    layout = l,
    vertex.size = sqrt(deg),
    vertex.label = V(g)$media,
    edge.arrow.size = .4,
    edge.curved = 0.5
)

```

There are easily separable groups of vertices.

# Discussion

In this we could easily see the "fat tails" in the graphs - vertices that "connects" another vertices that are less
"popular". Additionally, these graphs could be divided into separate graphs and this would make perfect sense (e.g. dividing into
families in case of GoT).
