---
title: "Social Networks and System Recomendations"
subtitle: "Project 4"
author: "Bartosz Rożek"
date: last-modified
date-format: "DD-MM-YYYY"
editor: visual
title-block-banner: "#e8e6ef"
page-layout: article
css: ../Styles/styles.css
format: 
  html:
    toc: true
    smooth-scroll: true
    embed-resources: true
    fontcolor: black
    toc-location: right
    toc-title: Summary
    toc-depth: 3
    number-sections: true
    number-depth: 4
    code-fold: false
    code-summary: "Code"
tbl-cap-location: bottom
---

```{r message=F}
library(ggplot2)
library(igraph)
library(purrr)
library(data.table)
```

# Exercise 0.1

```{r}
ER_adj_matrix <- function(N, p){
    mat <- as.numeric(runif(N^2) < p)
    dim(mat) <- c(N, N)
    mat[lower.tri(mat)] <- mat[upper.tri(mat)]
    diag(mat) <- 0
    return(mat)
}
```

# Exercise 0.2

```{r}
ad_mat <- ER_adj_matrix(25, 0.3)
graph_from_adjacency_matrix(ad_mat,
                            mode = "undirected") |> plot()
```

# Exercise 0.3

```{r}
colSums(ad_mat) |> hist()
```

# Exercise 0.4

We would expect binomial distribution since for every vertex we "draw" number of edges.

# Exercise 0.5

"The Poisson distribution is actually a limiting case of a Binomial distribution when the number of trials, n, gets very large and p, the probability of success, is small. As a rule of thumb, if $n \geq 100$ and $np\leq10$, the Poisson distribution (taking $\lambda=np$) can provide a very good approximation to the binomial distribution." - [Source](https://mathcenter.oxford.emory.edu/site/math117/connectingPoissonAndBinomial/)

# Exercise 0.6 & 0.7

```{r message=FALSE}

test_distributions <- function(vertex_dist, approx_dist){
    comp_df <- data.frame(values = c(vertex_dist, approx_dist), 
                      group = c(rep("Actual", n), rep("Poisson", n)))

    ks_pval <- suppressWarnings(ks.test(vertex_dist, approx_dist)$p.value)
    chi_pval <- suppressWarnings(chisq.test(vertex_dist, approx_dist)$p.value)
    g <- ggplot(comp_df, aes(values, fill = group)) +
        geom_histogram(position = "dodge") + 
        ggtitle(glue::glue("KS test pvalue: {round(ks_pval, 5)}\nChi-squared test pvalue: {round(chi_pval,5)}"))
    return(g)

}

ns <- c(10, 50, 100, 300)
ps <- c(0.01, 0.02, 0.1)
for (n in ns) {
    for (p in ps) {
        am <- ER_adj_matrix(n, p)
        vertex_dist <- colSums(am)
        approx_dist <- rpois(n, n*p)
        tryCatch(
        {
            test_distributions(vertex_dist, approx_dist) |> print()
        },
        error = function(cond) {
            print(glue::glue("Error for n={n}, p={p}: {cond$message}"))
            # Choose a return value in case of error
        },
        warning = function(cond) {
            print(cond)
        }
    )
    }
}
```

# Exercise 1

$$
q := \langle k \rangle
$$

$$ \mathbb{E}(K) = \sum_{k=0}^\inf \frac{ke^{-q}q^k }{k!}=
    \sum_{k=0}^\inf \frac{e^{-q}q^k }{(k-1)!} = 
    e^{-q}q \sum_{k=0}^\inf \frac{q^{k-1} }{(k-1)!} = 
    e^{-q}qe^{q} = q = \langle k \rangle
$$

\begin{aligned}
\mathbb{E}(K^2) = \sum_{k=0}^\inf \frac{k^2e^{-q}q^k }{k!}=
                  \sum_{k=0}^\inf \frac{ke^{-q}q^k }{(k-1)!}=
                  \sum_{k=0}^\inf \frac{(k - 1)e^{-q}q^k + e^{-q}q^k}{(k-1)!}= \\
                  e^{-q}q^2\sum_{k=0}^\inf \frac{q^{k-2} }{(k-2)!} + \sum_{k=0}^\inf \frac{e^{-q}q^k }{(k-1)!} = 
                  e^{-q}q^2e^{-q} + q = q^2 + q = \langle k \rangle + \langle k \rangle^2
\end{aligned}

# Exercise 2 

```{r}
cluster_coeff <- function(am) {
    ks <- rowSums(am)
    map_dbl(1:n, \(i) {
        neighbors <- which(am[i, ] == 1)
        if (length(neighbors) < 2){
            0
        }else{
             pairs <- combn(neighbors, m = 2, simplify = F)
            (map_int(pairs, \(pair) {
            am[pair[1], pair[2]]
        }) |> sum() * 2) / (ks[i] * (ks[i] - 1))
        }
    }) |>  mean()
    
} 

n <- 10
p <- 0.5
am <- ER_adj_matrix(n, p)
ks <- rowSums(am)
i <- 3
neighbors <- which(am[i,] == 1)
pairs <- combn(neighbors, m = 2, simplify = F)


map_dbl(1:100, \(j){
    am <- ER_adj_matrix(n, p)
    cluster_coeff(am)
}) |> boxplot()

```

# Exercise 3

```{r}

community_graph <- function(N, ns, P, ps){
    matrices <- map2(ns, ps, \(n,p) ER_adj_matrix(n,p))

    agg_mat <- ER_adj_matrix(N,P) 
    idxs <- cumsum(ns)
    
    pwalk(list(c(1,idxs[-length(ns)] + 1), 
                idxs,
                matrices), 
                \(idx_start, idx_stop, matrix_){
                    agg_mat[idx_start:idx_stop, idx_start:idx_stop] <<- matrix_
                })
    
    return(agg_mat)
}
N <- 20
P <- 0.03
ns <- c(5,5,7,3)
ps <- c(0.2, 0.12, 0.28, 0.3)
cg <- community_graph(N, ns, P, ps)
graph_from_adjacency_matrix(cg,
                            mode = "undirected") |> plot()

for (p_ in c(0.05, 0.1, 0.2,0.5)){
    ns <- c(5,5,5,5)
    ps <- rep(p_, 4)
    cg <- community_graph(N, ns, P, ps)
    graph_from_adjacency_matrix(cg,
                                mode = "undirected") |> plot()
}
```

# Exercise 4

```{r cache=TRUE}

WS_adj_matrix <- function(N, p, K){
    K <- 2
    k <- K/2
    mat <- matrix(0, N, N)

    for (i in 1:N) {
        left_neighboors <- seq(i + 1, i + k) %% N
        right_neighboors <- seq(i - 1, i - k) %% N
        neighboors <- c(left_neighboors, right_neighboors)
        neighboors[neighboors == 0] <- N
        mat[i, neighboors] <- 1
        mat[neighboors, i] <- 1
    }
    
    for (i in 1:N){
        edges <- which(mat[i,] == 1)
        non_edges <- which(mat[i,] == 0)
        for (edge in edges) {
            if (runif(1) < p){
                mat[i, edge] <- 0
                mat[edge, i] <- 0
                new_connection <- sample(non_edges, 1)
                mat[i, new_connection] <- 1
                mat[new_connection, i] <- 1
                non_edges <- non_edges[non_edges != new_connection]
                non_edges <- c(non_edges, edge)
            }
        }
    }
    
    return(mat)
}

ws_results <- map(seq(0.01, 0.7, 0.01), \(p) {
    avg_cc <- map_dbl(1:100, \(i) {
        WS_adj_matrix(20, p, 2) |> cluster_coeff()
    }) |> mean()
    return(data.frame(p = p, avg_cc = avg_cc))
}) |> rbindlist()

ggplot(ws_results, aes(p, avg_cc)) +
    geom_point() +
    geom_smooth(formula = "y~x", method = "loess")

```

# Exercise 6

```{r}

vectorInList <- function(v, L){
    idx <- map_lgl(L, \(elem){
        all(sort(elem) == sort(v))
    }) |> which()
    if (identical(idx, integer(0))) {
        idx <- 0
    }
    return(idx)
}

drop_zero_vertixes <- function(value, L, degrees) {
    # value <- 1
    # L <- possible_pairs
    if (degrees[value] == 0) {
        idxs <- map_lgl(L, \(elem) {
            value %in% elem
        }) |> which()
        L[idxs] <- NULL
    }
        return(L)
    }



configuration_model_am <- function(degrees) {
    if (!sum(degrees) %% 2) {
        stop("Sum must be even!")
    }
    orig_degrees <- degrees
    
    vertixes <- which(degrees > 0)
    possible_pairs <- combn(vertixes, m = 2, simplify = F)
    N <- length(degrees)
    mat <- matrix(0, N, N)
    
    while (length(possible_pairs) > 0) {
        if (sum(degrees > 0) == 1) {
            verts <- which(degrees > 0)
            mat[verts, verts] <- mat[verts, verts] + 1
        } else{
            verts <- sample(which(degrees > 0), 2)
            if (length(possible_pairs) > 0 &&
                !vectorInList(verts, possible_pairs)) {
                next
            }
            mat[verts[1], verts[2]] <- mat[verts[1], verts[2]] + 1
            mat[verts[2], verts[1]] <- mat[verts[2], verts[1]] + 1
            possible_pairs[vectorInList(verts, possible_pairs)] <- NULL
            
        }
        
        degrees[verts] <- degrees[verts] - 1
        possible_pairs <-
            drop_zero_vertixes(verts[1], possible_pairs, degrees)
        possible_pairs <-
            drop_zero_vertixes(verts[2], possible_pairs, degrees)
        
    }
    coef <- sum(degrees)/sum(orig_degrees)
    while (sum(degrees > 0) > 0) {
        if (sum(degrees > 0) == 1) {
            verts <- which(degrees > 0)
            mat[verts, verts] <- mat[verts, verts] + 1
        } else{
            verts <- sample(which(degrees > 0), 2)
            mat[verts[1], verts[2]] <- mat[verts[1], verts[2]] + 1
            mat[verts[2], verts[1]] <- mat[verts[2], verts[1]] + 1
            
        }
        degrees[verts] <- degrees[verts] - 1
        
    }
    return(list(mat = mat, coef = coef))
}
```

```{r cache=TRUE}

results <- map(seq(0.01, 0.7, 0.05), \(p){
    degrees <- rpois(25, p*100)
    if (!sum(degrees) %% 2){
        degrees[1] <- degrees[1] + 1
    }
    cm <- configuration_model_am(degrees)
    cm
})
results_df <- data.frame(p = seq(0.01, 0.7, 0.05),
    coef = map_dbl(results, \(el) el$coef))
ggplot(results_df, aes(p, coef)) +
    geom_point() +
    geom_smooth(formula = "y~x", method = "loess") +
    ggtitle("Probability test")
```

```{r cache=TRUE, warning=FALSE}
results <- map(seq(20, 100, 20), \(n){
    degrees <- rpois(n, n*0.1)
    if (!sum(degrees) %% 2) {
        degrees[1] <- degrees[1] + 1
    }
    cm <- configuration_model_am(degrees)
    cm
})
results_df <- data.frame(p = seq(20, 100, 20),
    coef = map_dbl(results, \(el) el$coef))
ggplot(results_df, aes(p, coef)) +
    geom_point() +
    geom_smooth(formula = "y~x", method = "loess") +
    ggtitle("Number test")

```